---
title: "US_Accidents"
author: "Vijeta Mishra"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
 html_document:
  theme: flatly
toc: TRUE
toc_float: TRUE
code_download: TRUE
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE, cache = F}
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,
  warning= FALSE,
  message= FALSE)
```
Introduction- 
Road accident has become very common these days. In USA more than 38,000 people die in road crash each year and on an average 4.4 million people are injured seriously enough to require medical attention. Source- https://www.asirt.org/safe-travel/road-safety-facts/ 

Looking at the severity of the accidents, for this project we have decided to use the US Accident- A Countrywide Traffic Accident Dataset (2016-2000) data-set to find something helpful.
This is a country wide traffic accident data-set which covers 49 states of the US. There are about 4.2 million accident records in this data-set.It contains all sort of information related to each accident like the weather condition during the accident, which side the accident occurred, address, time of accident etc. There are a total of 49 observations.

Below mentioned are the variables description-


```{r}

#run library

library(tidyverse)

#import dataset

US_Accidents <- read_csv("US_Accidents.csv")

#View data-set in new window

View(US_Accidents)

#Check column names of data-set

colnames(US_Accidents)

#Check number of rows present in data-set

nrow(US_Accidents)

#Sampling US_Accidents data-set. The new data-set will contain 1Million Row

US_Accidents_Sample <- US_Accidents[sample(nrow(US_Accidents), 1000000, replace = FALSE, prob = NULL),]


#Checking type of data frame to confirm its tibble
str(US_Accidents_Sample)

#View sample dataset in new window and check column names and datatype

View(US_Accidents_Sample)# to view compelete dataset.

glimpse(US_Accidents_Sample) # to check datatype of each column

colnames(US_Accidents_Sample) # to check column names
 
#View Unique State in Sample data

unique(US_Accidents_Sample$State)
 
#Export sample data to share with Team members

write_csv(US_Accidents_Sample,"US_Accidents_Sample.csv")
 

#Tidy data-set- This data has 1 million rows and it has many anomalies like the date and time format is not correct, there are many null values in the data-set etc. To address all these tidying data is an important step to be followed. Below mentioned are codes to tidy our data set.

#Separating the date and time from column Start time and end time.

US_Accidents_Sample <- US_Accidents_Sample %>% 
  separate(Start_Time, into= c("Accident_year","Start_Accident_month", "Start_Accident_date", "Start_Accident_Hour","Start_Accident_min","Start_Accident_sec")) %>% separate(End_Time, into= c("End_Accident_year","End_Accident_month", "End_Accident_date", "End_Accident_Hour","End_Accident_min","End_Accident_sec"))


#Removing columns which are not required for our analysis, to know the column names or number, we first check the column name and then use subset and select function to remove the unwanted columns. We also have the new dataset with the same name to avoid confusion

colnames(US_Accidents_Sample)

US_Accidents_Sample <- US_Accidents_Sample %>% 
select(-c(2,3,19,20,29,30,32,33,55,57,58,59))

#Accident Per Year- This graph shows the increase in number of accidents with increase in year.

by_year<- US_Accidents_Sample %>% 
  group_by(Accident_year) %>% 
  summarise(total_accident= n())

ggplot(data = US_Accidents_Sample)+
  geom_bar(mapping = aes(x= Accident_year, color= Severity))+
 ggtitle("Accidents by Year")+
  theme_grey()

#* not able to color severity

#Accident per year with severity shows that while there is an overall increase in number of accidents every passing year and 2020 has the highest number of accidents. However, severity 3 has decreased in year 2020

ggplot(data = US_Accidents_Sample)+
  geom_bar(mapping = aes(x= Accident_year,colour = "Severity"), colour = "blue", fill = "blue")+
  ggtitle("Accidents by Year")+
  theme_grey()+
  facet_wrap(~Severity)


#Accident by State, arranged in descending order

by_state<- US_Accidents_Sample %>% 
  group_by(State) %>% 
  summarise(No.of_accident = n()) %>% 
  arrange(desc(No.of_accident))
 
#Summary of accidents by state 

summary(by_state)

# graph showing top 5 state that has highest number of accidents.

top_states<- top_n(by_state, 5)

ggplot(data= top_states, aes(x=State, y=No.of_accident)) +
  geom_histogram(stat="identity", fill="steelblue")+
   ggtitle("Top 5 Accident States")+
  geom_text(aes(label=No.of_accident), vjust=1.5, color="white", size=3.5)+
  theme_classic()

# Accidents by timezone
by_timezone <- US_Accidents_Sample %>% 
  group_by(Timezone) %>% 
  summarise(accident_by_timezone= n())


#Accident Severity Count
ggplot(data= US_Accidents_Sample, mapping = aes(x= Severity))+
  geom_bar(mapping = aes(fill = Sunrise_Sunset))+
  ggtitle("Accident Severity Count Barchart")

#Analyze which side of the road has more accidents.

by_side <- US_Accidents_Sample %>% 
  group_by(Side) %>% 
  summarise(count = n())

  ggplot(data = by_side, mapping = aes (x= Side,y= count))+
  geom_bar(stat="identity")+
    ggtitle("Accidents_by_Side")+
    geom_text(aes(label= count), vjust=1.5, color="white", size=3.5)+
  theme_classic()

#Frequency of accidents happening in different time of the day.
by_timeofDay<-
  US_Accidents_Sample %>% 
  group_by(Start_Accident_Hour) %>% 
  summarise(accident_count= n())

ggplot(data = US_Accidents_Sample)+
  geom_bar(mapping = aes(x=Start_Accident_Hour, shade = Severity))

```


